Transfer of Learned Opponent Models in Zero Sum Games

### 2020-01-15 Report/draft_report_v1.tex
finalise v1 of report

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Transfer of Learned Opponent Models in Zero Sum Games},
            pdfauthor={Ismail Guennouni~\& Maarten Speekenbrink},
            pdfkeywords={keywords},
            colorlinks=true,
            linkcolor=blue,
            filecolor=Maroon,
            citecolor=Blue,
            urlcolor=Blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\shorttitle{draft}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Experimental Psychology, University College London}
\keywords{keywords\newline\indent Word count: X}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\usepackage{lineno}

\linenumbers
\usepackage[document]{ragged2e}

@@

\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

@@

In order to explore learning transfer in strategic settings, it is generally useful to study simple games as a model of more complex interactions. More specifically, when studying learning, we need a framework that allows the study of whether and how a player takes into consideration, over time, the impact of its current and future actions on the future actions of the opponent and the future cumulative rewards. Repeated games, in which players interact repeatedly with the same opponent and have the ability to learn about the opponent's strategies and preferences (Mertens, 1990) are particularly adapted to modelling learning about the opponent.

Early literature on learning transfer in games has mostly focused on measuring the proportion of people who play normatively optimal (Nash Equilibria) or salient actions (e.g Risk Dominance) in later games, having had experience with a similar game environment previously. For instance, Ho et al. (1998) measure transfer as the proportion of players who choose the Nash Equilibrium in later p-beauty contest games, after traingn on similar games. They find there is no evidence of immediate transfer (Nash equilibrium play in the first round of the new game) but positive structural learning transfer as shown by the faster convergence to equilibrium play by experienced vs non experienced players. Camerer \& Knez (2000) test learning transfer in players exposed to two games with multiple equilibria sequentially and explore the ability of players to conordinate their actions to choose a particular equilibrium in subsequent games having reached it in prior ones. They distinguish between games that are similar in a purely descriptive way, meaning similar choice labels, identity of players, format and number of action choices; and games that are similar in a strategic sense, meaning similar payoffs from combination of actions, identical equilibrium properties or significant social characteristics of payoffs such as possibility of punishment, need for fairness and cooperative vs competitive settings. They find that transfer of learning (successful coordination) occurs more readily in the presence of both descriptive and strategic similarity. If the games were only strategically similar, then the transfer was much weaker.

@@

One of the major limitations in studies of how humans adapt to computerised opponents is that they have mostly looked at the ability of players to detect and exploit action-based learning rules. They either used deterministic strategies that did not consider the human participant's play, as in the case of pre-determined mixed strategy algorithms, or strategies that only depended on the player's patterns of past plays. These strategies did not necessarily take into account the computer's own past plays, or the degree of strategic reasoning of the human player about the computer. Their play was not \enquote{human-like} in the sense that humans are not very good at playing specific mixed strategies with any precision of at detecting patterns from long sequences of past play due to cognitive constraints. It is therefore important to have agents that \enquote{play like humans}, and one way of achieving that is to embed theses agents with human-like theory of mind abilities based on limited steps of recursive reasoning. Simon (1972) explains that humans have limited cognitive capacities and as such cannot be expected to solve computationally intractable problems such as finding Nash equilibria. Instead, they will try to \enquote{satisfice} by choosing a strategy that is adequate in a simplified model of the environment, rather than an optimal one. This concept finds its natural application in \enquote{level-k} theory, first adopted by Stahl \& Wilson (1995). It posits that deviations from Nash equilibrium solutions in simple games are explained by the fact that humans have a heterogenous degree of strategic sophistication. At the bottom of the ladder, level-0 players are non-strategic and play either randomly or use a salient strategy in the game environment (Arad \& Rubinstein, 2012). Level-1 players are next up the ladder of strategic sophistication and will assume all their opponents belong to the level-0 category and as such will best respond to them given this assumption. Likewise, a level-2 player will choose actions that are the best response given the belief that all opponents are exactly one level below and so on.

@@

Participants were informed they would play three different games against the same computer opponent. Each participant plays all three games consecutively and in the same order described above. Participants were told that the opponent cannot cheat and will choose its actions simultaneously without knowledge of the participant's choice. A total of 50 rounds of each game was played with the player's score displayed at the end of each game. The score was calculated as the number of wins minus the number of losses. Ties did not affect the score. In order to incentivise the participants to maximise the number of wins against the opponents, players were paid a bonus at the end of the experiment that was proportional to their final score. The overall score of the players was translated into the bonus by making each point worth \$£\$0.02. This bonus is significant as players could increase the total payoff from the experiment by up to 60\% assuming they'd won all rounds against the computer opponent.

@@

In the second experiment, participants each played 3 games sequentially against both level-1 and level-2 computer opponents, rather than just one like in the first experiment. The three games were Rock-Paper-Scissors, Fire-Water-Grass, and the penalty shootout game. The first two games were identical to the ones used in the first experiment. In the final game (shootout) the participants played the role of the player shooting a football (soccer) penalty, and the AI opponent was the goalkeeper. Players had the choice between three actions, like in the first two games: Shooting the football to the left, right or centre of the goal. If the player shoots in a direction different from that of where the goalkeeper dives, they win the round and the goalkeeper loses. Otherwise, the goalkeeper catches the ball and the player loses the round. There is no possibility of ties in this game. Figure XXXXXXXXX shows a snapshot of play in the shootout game. What makes this game different however is that there are two ways to beat the opponent in each round: if we think the opponent is going to choose \enquote{'right}' in the next round, we can win by both choosing \enquote{'left}' and \enquote{'center}'. A level-1 player (thinks that his opponent will repeat his last action) has two ways to win the next round. As such, we have programmed the level-2 computer program to choose randomly between the two possibilities that a level-1 player may choose.

@@

In order to encourage participants to think about their next choice, a countdown timer of 3 seconds was introduced at the beginning of each round. During those 3 seconds, participants could not choose an option and had to wait for the timer to run out. A small delay that changed randomly (between 0.5 and 3 seconds) was also introduced in the time it took the AI agent to give back their response to make it seem like it is also thinking about its last actions, as a way of simulating a real human opponent. After each round, the participants were given detailed feedback about their opponent actions as well as whether they won or lost the last round. Further information about the outcome of previous rounds was also visible on the game screen below the feedback area, throughout each stage game and opponent could go back many rounds to study the history of interaction. The number of wins, losses and ties were clearly shown at the top of the screen for each game, and this scoreboard was reinitialised to zero at the onset of a new stage game. An example of the interface for the rock-paper-scissors game is provided in Figure XX.

@@

To analyse within and between game learning, we used a 2 (condition: level-1, level-2) by 3 (game: RPS, FWG, NUMBERS) by 2 (block: first half, second half) repeated measures ANOVA with the first factor varying between participants. As the number of participants facing level-2 agents was slightly higher than those facing level-1 agents (28 vs 24), the ANOVA is unbalanced. We therefore used the \enquote{afex} (Singmann \& al., 2016) package in R, as it allows us to handle unbalanced designs and produces type III sum of squares by default.

@@

As a measure for learning transfer, we focus on participants' scores in the earliest rounds, excluding the very first round for which the agent is programmed to play randomly as it has no data on prior rounds on which to build its response. A group of players with no experience of the game are expected to have scores not significantly different from 0. Any significantly positive group average scores would therefore reflect prior learning from past experiences. In figure XXX, the average score across participants by game for rounds 2-6 are plotted. Scores are also averaged across levels of condition. We test the average scores for each game against a hypothesised value of 0 for a non-experienced player using parametric one sample t-tests. As expected for the initial RPS game, the average score is not significantly different from 0 ( t(148.85) = 1.04 , p = 0.89). In FWG, the score is significantly higher than 0 ( t(148.85) = 4.58 , p \textless{} 0.0001). Unexpectedly, this is also the case for NUMBERS ( t(148.85) = 3.00, p = 0.0092).

\hypertarget{moderation-of-learning-transfer-by-type-of-opponent-and-game-similarity}{%
\subsubsection{Moderation of learning transfer by type of opponent and game similarity}\label{moderation-of-learning-transfer-by-type-of-opponent-and-game-similarity}}

Figure 10 shows the marginal mean scores for each game for both level-1 and level-2 facing players. Graphically we can see that the pattern is dissimilar between level-1 and level-2 players, and we suspect transfer to be positively associated with similarity and negatively with degree of sophistication of the agent. To test these hypotheses, we run statistical tests on scores by game and opponent against the null hypothesis of 0 (no transfer). Table XXXXXXX shows the scores on rounds 2-6 for both type of opponents in each game. We also report the t-stats and p-values to test the null hypothesis that scores on these early rounds are equal to 0. Rejecting the null hypothesis means that prior learning has been used to beat the opponent in these early rounds. For level-1 facing players, there is evidence of learning transfer to both FWG and NUMBERS. For level-2 facing players, there is evidence for transfer to the similar game FWG (albeit scores are lower than for level-1 player) but not to the dissimilar game of NUMBERS.

TABLE OF SCORES BY OPPONENT AND GAME (PG 40)

Our results when averaging across conditions (previous section) showed that there was indeed evidence for transfer to the more dissimilar game (NUMBERS). We can see from TABLE XXXXX that this transfer is exclusively driven by level-1 facing players, as scores of level-2 facing players are close to nil. Therefore, both groups of condition can generalise to the similar game, but only those facing the less sophisticated opponent are able to generalise to the less similar game.

@@

Using the average scores obtained by participants in each game and interaction, we explore whether learning has occurred within and between games. We perform a two (condition: level-1 first, level-2 first) by two ( opp\_type: level\_1 or level\_2) by three (game: RPS, FWG, SHOOT) by two (interaction: first or second) repeated measures ANOVA with the first factor varying between participants. As the number of participants facing level-1 agents was slightly higher than those facing level-2 agents (27 vs 23), the ANOVA will not be balanced between the two levels of condition. We therefore used the \enquote{afex} (Singmann \& al., 2016) package in R, as it allows us to handle unbalanced designs and produces type III sum of squares by default. The full results of this analysis are provided in Table XXXXXXXXX.

There is evidence for a main effect of Game on scores (F(1.85,88.7) = 11.81, ges = 0.04, p \textless{} .0001), meaning that we reject the null hypothesis of scores in games being equal. To explore these differences further, we look at post-hoc analyses for pairwise comparisons between game scores (p-values adjusted using Holm method for multiple comparisons). We find the performance in the games increases steadily throughout the experiment, with FWG performance significantly higher than RPS (t(96) =2.53, p = 0.025), and performance in the SHOOT game also significantly higher than in FWG ( t(96) = 2.32, p = 0.025 ). There was no main effect of either opponent type, the interaction factor( first or second time opponent was faced) , or the condition factor (whether level-1 or level-2 opponent was faced first). This means that if we ignore potential secondary effects from interactions, neither the type of opponent faced, nor whether it was the first or second interaction with the opponent nor which opponent was faced first did have a main effect on performance.

@@

As a measure for learning transfer we will compare scores only on rounds 2-6 of each game (excluding the very first round where play is necessarily random). Because the number of rounds here is very limited, there should be very little learning within games, and we should be better measure learning transfer if it exits. In the figure below, we show the average score across participants in rounds 2-6. These scores are also averaged across the levels of condition (meaning they are irrespective of which opponent players faced first). For both the FWG and SHOOT games, score in the early rounds of the first stage are significantly higher than 0 for both opponent types. (Level-1 opponent: FWG: t(270) = 4.99, p \textless{} 0.0001; SHOOT: t(270) = 6.66, p \textless{} 0.0001; Level-2 opponent: FWG: t(270) = 4.40, p \textless{} 0.0001; SHOOT: t(270) = 3.21, p =0.004 ). Looking more specifically at early scores by type of opponent faced, we confirm the result from the first experiment that it is easier to transfer learning to the more dissimilar game (Shoot) when facing a level 1 opponent. Indeed, while the early scores of FWG for level-1 and level-2 facing players are not significantly different from each other, the score of the players facing the level-1 opponent is indeed almost 0.2 point per round higher than that of players facing level-2 opponents, and the difference is statistically significant ( t(144) = 2.45 , p = 0.01). These early scores have also been adjusted to account for the fact that the shootout game has higher average scores when playing randomly, and therefore this difference is really due to better learning transfer and not due to chance.

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Transfer of Learned Opponent Models in Zero Sum Games},
            pdfauthor={Ismail Guennouni~\& Maarten Speekenbrink},
            pdfkeywords={keywords},
            colorlinks=true,
            linkcolor=blue,
            filecolor=Maroon,
            citecolor=Blue,
            urlcolor=Blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother

@@

\authornote{

@@

In order to explore learning transfer in strategic settings, it is generally useful to study simple games as a model of more complex interactions. More specifically, we need a framework that allows the study of whether and how a player takes into consideration, over time, the impact of its current and future actions on the future actions of the opponent and the future cumulative rewards. Repeated games, in which players interact repeatedly with the same opponent and have the ability to learn about the opponent's strategies and preferences (Mertens, 1990) are particularly adapted to modelling learning about the opponent.

Early literature on learning transfer in games has mostly focused on measuring the proportion of people who play normatively optimal (Nash Equilibria) or salient actions (e.g Risk Dominance) in later games, having had experience with a similar game environment previously. For instance, Ho et al. (1998) measure transfer as the proportion of players who choose the Nash Equilibrium in later p-beauty contest games, after training on similar games. They find there is no evidence of immediate transfer (Nash equilibrium play in the first round of the new game) but positive structural learning transfer as shown by the faster convergence to equilibrium play by experienced vs non experienced players. Camerer \& Knez (2000) test learning transfer in players exposed to two games with multiple equilibria sequentially and explore the ability of players to coordinate their actions to choose a particular equilibrium in subsequent games having reached it in prior ones. They distinguish between games that are similar in a purely descriptive way, meaning similar choice labels, identity of players, format and number of action choices; and games that are similar in a strategic sense, meaning similar payoffs from combination of actions, identical equilibrium properties or significant social characteristics of payoffs such as possibility of punishment, need for fairness and cooperative vs competitive settings. They find that transfer of learning (successful coordination) occurs more readily in the presence of both descriptive and strategic similarity. If the games were only strategically similar, then the transfer was much weaker.

@@

Still, one of the limitations in studies of how humans adapt to computerised opponents is that they have mostly looked at the ability of players to detect and exploit action-based learning rules. They either used deterministic strategies that did not consider the human participant's play, as in the case of pre-determined mixed strategy algorithms, or strategies that only depended on the player's patterns of past plays. These strategies did not necessarily take into account the computer's own past plays, or the degree of strategic reasoning of the human player about the computer. Their play was not ``human-like'' in the sense that humans are not very good at playing specific mixed strategies with any precision of at detecting patterns from long sequences of past play due to cognitive constraints. It is therefore important to have agents that ``play like humans'', and one way of achieving that is to embed theses agents with human-like theory of mind abilities based on limited steps of recursive reasoning. Simon (1972) explains that humans have limited cognitive capacities and as such cannot be expected to solve computationally intractable problems such as finding Nash equilibria. Instead, they will try to `satisfice' by choosing a strategy that is adequate in a simplified model of the environment, rather than an optimal one. This concept finds its natural application in `level-k' theory, first adopted by Stahl \& Wilson (1995). It posits that deviations from Nash equilibrium solutions in simple games are explained by the fact that humans have a heterogenous degree of strategic sophistication. At the bottom of the ladder, level-0 players are non-strategic and play either randomly or use a salient strategy in the game environment (Arad \& Rubinstein, 2012). Level-1 players are next up the ladder of strategic sophistication and will assume all their opponents belong to the level-0 category and as such will best respond to them given this assumption. Likewise, a level-2 player will choose actions that are the best response given the belief that all opponents are exactly one level below and so on.

@@

Participants were informed they would play three different games against the same computer opponent. Each participant plays all three games consecutively and in the same order described above. Participants were told that the opponent cannot cheat and will choose its actions simultaneously without knowledge of the participant's choice. A total of 50 rounds of each game was played with the player's score displayed at the end of each game. The score was calculated as the number of wins minus the number of losses. Ties did not affect the score. In order to incentivise the participants to maximise the number of wins against the opponents, players were paid a bonus at the end of the experiment that was proportional to their final score. The overall score of the players was translated into the bonus by making each point worth £0.02. This bonus is significant as players could increase the total payoff from the experiment by up to 60\% assuming they'd won all rounds against the computer opponent.

@@

In the second experiment, participants each played 3 games sequentially against both level-1 and level-2 computer opponents, rather than just one like in the first experiment. The three games were Rock-Paper-Scissors, Fire-Water-Grass, and the penalty shootout game. The first two games were identical to the ones used in the first experiment. In the final game (shootout) the participants played the role of the player shooting a football (soccer) penalty, and the AI opponent was the goalkeeper. Players had the choice between three actions, like in the first two games: Shooting the football to the left, right or centre of the goal. If the player shoots in a direction different from that of where the goalkeeper dives, they win the round and the goalkeeper loses. Otherwise, the goalkeeper catches the ball and the player loses the round. There is no possibility of ties in this game. Figure XXXXXXXXX shows a snapshot of play in the shootout game. What makes this game different however is that there are two ways to beat the opponent in each round: if we think the opponent is going to choose `'right'' in the next round, we can win by both choosing `'left'' and `'center''. A level-1 player (thinks that his opponent will repeat his last action) has two ways to win the next round. As such, we have programmed the level-2 computer program to choose randomly between the two possibilities that a level-1 player may choose.

@@

In order to encourage participants to think about their next choice, a countdown timer of 3 seconds was introduced at the beginning of each round. During those 3 seconds, participants could not choose an option and had to wait for the timer to run out. A small delay that changed randomly (between 0.5 and 3 seconds) was also introduced in the time it took the AI agent to give back their response to make it seem like it is also thinking about its last actions, as a way of simulating a real human opponent. After each round, the participants were given detailed feedback about their opponent actions as well as whether they won or lost the last round. Further information about the outcome of previous rounds was also visible on the game screen below the feedback area, throughout each stage game and opponent could go back many rounds to study the history of interaction. The number of wins, losses and ties were clearly shown at the top of the screen for each game, and this scoreboard was reinitialised to zero at the onset of a new stage game. An example of the interface for the rock-paper-scissors game is provided in Figure~\ref{fig:feedback-rps}.

\begin{figure}

{\centering \includegraphics{../Report/images/feedback_rps}

\caption{Screenshot of the feedback at the end of a round of Rock-Paper-Scissors}\label{fig:feedback-rps}
\end{figure}

@@

To analyse within and between game learning, we used a 2 (condition: level-1, level-2) by 3 (game: RPS, FWG, NUMBERS) by 2 (block: first half, second half) repeated measures ANOVA with the first factor varying between participants. As the number of participants facing level-2 agents was slightly higher than those facing level-1 agents (28 vs 24), the ANOVA is unbalanced. We therefore used the ``afex'' (Singmann \& al., 2016) package in R, as it allows us to handle unbalanced designs and produces type III sum of squares by default.

@@

As a measure for learning transfer, we focus on participants' scores in rounds 2-6, excluding the very first round for which the agent is programmed to play randomly as it has no data on prior rounds on which to build its response. A group of players with no experience of the game are expected to have scores not significantly different from 0. Any significantly positive group average scores would therefore reflect prior learning from past experiences. In Figure~@ref(fig: exp1-early-score), the average score across participants by game for rounds 2-6 are plotted. Scores are also averaged across levels of condition.

\begin{figure}

{\centering \includegraphics{../Report/images/exp1_early_scor_game}

\caption{Average scores for rounds 2-6 by game in first experiment}\label{fig:exp1-early-score}
\end{figure}

We test the average scores for each game against a hypothesised value of 0 for a non-experienced player using parametric one sample t-tests. As expected for the initial RPS game, the average score is not significantly different from 0 as this is the first game and no learning is possible (t(148.85) = 1.04 , p = 0.89). In FWG, the score is significantly higher than 0 ( t(148.85) = 4.58 , p \textless{} 0.0001). This is also the case for the more dissimilar game : NUMBERS ( t(148.85) = 3.00, p = 0.0092).

Next, we explore whether learning transfer is moderated by the type of opponent and game similarity. Figure~@ref(fig: exp1-score-by-opp) shows the marginal mean scores for each game for both level-1 and level-2 facing players. Graphically we can see that the pattern is dissimilar between level-1 and level-2 players, and we suspect transfer to be positively associated with similarity and negatively with degree of sophistication of the agent. To test these hypotheses, we run statistical tests on scores by game and opponent against the null hypothesis of 0 (no transfer).

\begin{figure}

{\centering \includegraphics{../Report/images/exp1_earlyScore_by_opp}

\caption{Marginal means of players scores by game and type of opponents}\label{fig:exp1-score-by-opp}
\end{figure}

Table~\ref{tab:table-early-scores} shows the scores on rounds 2-6 for both type of opponents in each game. We also report the t-stats and p-values to test the null hypothesis that scores on these early rounds are equal to 0. Rejecting the null hypothesis means that prior learning has been used to beat the opponent in these early rounds. For level-1 facing players, there is evidence of learning transfer from RPS to both FWG and NUMBERS. For level-2 facing players, there is evidence for transfer from RPS to the similar game FWG (albeit scores are lower than for level-1 player) but not to the dissimilar game of NUMBERS.

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:table-early-scores}Average early round scores by game and condition in first experiment}
\begin{tabular}{lllllllll}
\toprule
game.f & \multicolumn{1}{c}{condition} & \multicolumn{1}{c}{lsmean} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{df} & \multicolumn{1}{c}{lower.CL} & \multicolumn{1}{c}{upper.CL} & \multicolumn{1}{c}{t.ratio} & \multicolumn{1}{c}{p.value}\\
\midrule
FWG & Level1 & 0.30 & 0.08 & 149.58 & 0.10 & 0.51 & 3.96 & 0.00\\
NUM & Level1 & 0.29 & 0.08 & 149.58 & 0.08 & 0.49 & 3.74 & 0.00\\
RPS & Level1 & 0.03 & 0.08 & 149.58 & -0.18 & 0.23 & 0.38 & 1.00\\
FWG & Level2 & 0.18 & 0.07 & 147.52 & -0.01 & 0.38 & 2.48 & 0.06\\
NUM & Level2 & 0.03 & 0.07 & 147.52 & -0.16 & 0.23 & 0.43 & 1.00\\
RPS & Level2 & 0.08 & 0.07 & 147.52 & -0.11 & 0.28 & 1.11 & 0.80\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\end{table}

Our results when averaging across conditions (previous section) showed that there was indeed evidence for transfer to the more dissimilar game (NUMBERS). We can see from TABLE~\ref{tab:table-early-scores} that this transfer is exclusively driven by level-1 facing players, as scores of level-2 facing players are close to nil. Therefore, both groups of condition can generalise to the similar game, but only those facing the less sophisticated opponent are able to generalise to the less similar game.

@@

Using the average scores obtained by participants in each game and interaction, we explore whether learning has occurred within and between games. We perform a two (condition: level-1 first, level-2 first) by two ( opp\_type: level\_1 or level\_2) by three (game: RPS, FWG, SHOOT) by two (interaction: first or second) repeated measures ANOVA with the first factor varying between participants. As the number of participants facing level-1 agents was slightly higher than those facing level-2 agents (27 vs 23), the ANOVA will not be balanced between the two levels of condition. We therefore used the ``afex'' (Singmann \& al., 2016) package in R, as it allows us to handle unbalanced designs and produces type III sum of squares by default. The full results of this analysis are provided in Table~\ref{tab:exp2-anova}.

\begin{verbatim}
## Warning in summary.Anova.mlm(x$Anova): HF eps > 1 treated as 1
\end{verbatim}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp2-anova}ANOVA results for experiment 2}
\begin{tabular}{lllllll}
\toprule
Effect & \multicolumn{1}{c}{$F$} & \multicolumn{1}{c}{$\mathit{df}_1^{GG}$} & \multicolumn{1}{c}{$\mathit{df}_2^{GG}$} & \multicolumn{1}{c}{$\mathit{MSE}$} & \multicolumn{1}{c}{$p$} & \multicolumn{1}{c}{$\hat{\eta}^2_G$}\\
\midrule
Condition f & 0.45 & 1 & 48 & 0.43 & .505 & .003\\
Game f & 11.81 & 1.85 & 88.71 & 0.12 & < .001 & .039\\
Interaction lvl & 1.34 & 1 & 48 & 0.06 & .253 & .001\\
Opp type & 0.00 & 1 & 48 & 0.12 & .973 & .000\\
Condition f $\times$ Game f & 1.34 & 1.85 & 88.71 & 0.12 & .266 & .005\\
Condition f $\times$ Interaction lvl & 0.03 & 1 & 48 & 0.06 & .864 & .000\\
Condition f $\times$ Opp type & 0.13 & 1 & 48 & 0.12 & .717 & .000\\
Game f $\times$ Interaction lvl & 0.46 & 1.99 & 95.72 & 0.08 & .635 & .001\\
Game f $\times$ Opp type & 5.31 & 1.7 & 81.82 & 0.11 & .010 & .015\\
Interaction lvl $\times$ Opp type & 2.16 & 1 & 48 & 0.04 & .148 & .001\\
Condition f $\times$ Game f $\times$ Interaction lvl & 2.43 & 1.99 & 95.72 & 0.08 & .094 & .006\\
Condition f $\times$ Game f $\times$ Opp type & 1.23 & 1.7 & 81.82 & 0.11 & .294 & .003\\
Condition f $\times$ Interaction lvl $\times$ Opp type & 0.02 & 1 & 48 & 0.04 & .883 & .000\\
Game f $\times$ Interaction lvl $\times$ Opp type & 0.96 & 1.89 & 90.83 & 0.09 & .383 & .002\\
Condition f $\times$ Game f $\times$ Interaction lvl $\times$ Opp type & 0.94 & 1.89 & 90.83 & 0.09 & .390 & .002\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\end{table}

There is evidence for a main effect of Game on scores (F(1.85,88.7) = 11.81, ges = 0.04, p \textless{} .0001), meaning that we reject the null hypothesis of scores in games being equal. To explore these differences further, we look at post-hoc analyses for pairwise comparisons between game scores (p-values adjusted using Holm method for multiple comparisons). We find the performance in the games increases steadily throughout the experiment, with FWG performance significantly higher than RPS (t(96) =2.53, p = 0.025), and performance in the SHOOT game also significantly higher than in FWG ( t(96) = 2.32, p = 0.025 ). There was no main effect of either opponent type, the interaction factor( first or second time opponent was faced) , or the condition factor (whether level-1 or level-2 opponent was faced first). This means that neither the type of opponent faced, nor whether it was the first or second interaction with the opponent nor which opponent was faced first did have a main effect on performance.

@@

As a measure for learning transfer we will compare scores only on the first 5 rounds (rounds 2-6 of each game, excluding the very first round where play is necessarily random). Because the number of rounds here is very limited, there should be very little learning within games, and we should be better measure learning transfer if it exits.

\begin{figure}

{\centering \includegraphics{draft_report_v1_files/figure-latex/exp2-early-scores-1}

\caption{Average early round scores by game for experiment 2}\label{fig:exp2-early-scores}
\end{figure}

In Figure~\ref{fig:exp2-early-scores}, we show the average score across participants in rounds 2-6 of the first interaction with the opponent for each game. These scores are also averaged across the levels of condition (meaning they are irrespective of which opponent players faced first). For both the FWG and SHOOT games, score in the early rounds of the first interaction are significantly higher than 0 for both opponent types. (Level-1 opponent: FWG: t(270) = 4.99, p \textless{} 0.0001; SHOOT: t(270) = 6.66, p \textless{} 0.0001; Level-2 opponent: FWG: t(270) = 4.40, p \textless{} 0.0001; SHOOT: t(270) = 3.21, p =0.004 ). Looking more specifically at early scores by type of opponent faced, we confirm the result from the first experiment that it is easier to transfer learning to the more dissimilar game (Shoot) when facing a level 1 opponent. Indeed, while the early scores of FWG for level-1 and level-2 facing players are not significantly different from each other, the score of the players facing the level-1 opponent is indeed almost 0.2 point per round higher than that of players facing level-2 opponents, and the difference is statistically significant ( t(144) = 2.45 , p = 0.01). These early scores have also been adjusted to account for the fact that the shootout game has higher average scores when playing randomly, and therefore this difference is really due to better learning transfer and not due to chance.

\hypertarget{computational-modelling}{%
\section{Computational modelling}\label{computational-modelling}}

To gain more insight into how participants played the games against the computer opponents, we compared multiple models of strategies the players may have been using to learn how to beat the opponent. As a base model, we assume play is random, and each potential action is chosen with equal probability. Note that this corresponds to the Nash equilibrium strategy. Next, we include for comparison purposes a few other basic models of behavior, such as Q-learning on action choices, as well as a more sophisticatd Q-learning model incorporating a state space consisting of last round play. The Bayesian approach is represented by models whereby agents would update an initial belief over the opponent's strategy consisting of various level-k strategies (resp. 0,1 and 2). We fitted two models in which agents best respond to the opponent model. In the first, the beliefs about the opponent's adopted strategy are transfered across games, and another in which no transfer of opponent models is allowed.

Next, we fitted a self-tuning EWA model (Ho et al., 2004) both on action choices without states and with states consisting of lat round play. The self-tuning EWA model generalizes both reinforcement learning as well as fictitious play models and has been proven to do well than equilibrium models in various game situations. Finally, we also included a model-based reinforcement learning approach. whereby the agent learns the transition probabilities between states ans uses that to choose the best action, given the current state and it's belief about what the new state will be.
